Space is the Place: Effects of Continuous Spatial Structure on Analysis of Population Genetic Data

### 2019-12-05 ms/spaceness.tex
quick pass on the ms

In all simulations, the phenotype of each individual
is determined by drawing from a Gaussian distribution with standard deviation 10 \revpoint{1}{5}
and a mean that may depend on spatial position.
In spatially varying models, the mean phenotype differs by two standard deviations across the landscape.
We then adjust the geographic pattern of mean phenotype to create four types of spatially autocorrelated environmental influences on phenotype.
In the first simulation of \emph{nonspatial} environments, the mean did not change,
so that all individuals' phenotypes were drawn independently from a Gaussian distribution with mean 110 and standard deviation 10.
Next, to simulate \emph{clinal} environmental influences on phenotype, 
we increased the mean phenotype from 100 on the left edge of the range to 120 on the right edge (two phenotypic standard deviations).
Concretely, the mean phenotype $p$ for an individual at position $(x,y)$ is $p=100 + 2x / 5$. \revpoint{4}{10}
Third, we simulated a more concentrated ``\emph{corner}'' environmental effect by setting the mean phenotype to 120
for individuals with both $x$ and $y$ coordinates below 20 (two standard deviations above the rest of the map). 
Finally, in ``\emph{patchy}'' simulations we selected 10 random points on the map
and set the mean phenotype of all individuals within three map units of each of these points
to 120.

@@

Any simulation of a population of reproducing organisms for a reasonably long amount of time
requires some kind of control on population sizes,
or else the population will either die out or grow very large. \llabel{demog_disco}
The usual choice of population regulation for population genetics --
a constant size, as in the Wright--Fisher model --
implies biologically unrealistic interactions between geographically distant parts of the species range.
Our choice to regulate population size by including a local density-dependent control on mortality
is only one of many possible ways to do this.
We could have instead regulated fecundity, or recruitment, or both;
this general class of models
is sometimes referred to as the ``Bolker--Pacala model'' \citep{bolker1997using}.
It is not currently clear how much different choices of vital rates,
or of functional forms for the regulation,
might quantitatively affect our results,
although the general predictions should be robust to similar forms of regulation.
Since populations are still entirely \emph{intrinsically} regulated,
our model still has a very strong ``population genetics'' flavor.
Alternatively, population size could be regulated by interactions with other species
(e.g., a Lotka-Volterra model),
or extrinsically specified by local resource availability
(e.g., by food or nest site availability).
Indeed, our model could be interpreted as a caricature of such a model:
as local density increases, good habitat is increasingly occupied,
pushing individuals into more marginal habitat and increasing their mortality.
Many such models should be behave similarly to ours,
but others (especially those with local population cycling), may differ dramatically.
There is a clear need for greater exploration of 
the consequences for population genetics of ecologically realistic population models.

>>>>

In all simulations, the phenotype of each individual
is determined by drawing from a Gaussian distribution with standard deviation 10 \revpoint{1}{5}
and a mean that depends on spatial position.
In spatially varying models, the mean phenotype differs by two standard deviations across the landscape.
We then adjust the geographic pattern of mean phenotype to create four types of spatially autocorrelated environmental influences on phenotype.
In the first simulation of \emph{nonspatial} environments, the mean did not change,
so that all individuals' phenotypes were drawn independently from a Gaussian distribution with mean 110 and standard deviation 10.
Next, to simulate \emph{clinal} environmental influences on phenotype, 
we increased the mean phenotype from 100 on the left edge of the range to 120 on the right edge (two phenotypic standard deviations).
Concretely, the mean phenotype $p$ for an individual at position $(x,y)$ is $p=100 + 2x / 5$. \revpoint{4}{10}
Third, we simulated a more concentrated ``\emph{corner}'' environmental effect by setting the mean phenotype to 120
for individuals with both $x$ and $y$ coordinates below 20 (two standard deviations above the rest of the map). 
Finally, in ``\emph{patchy}'' simulations we selected 10 random points on the map
and set the mean phenotype of all individuals within three map units of each of these points
to 120.

@@

Any simulation of a population of reproducing organisms
requires some kind of control on population sizes,
or else the population will either die out or grow very large. \llabel{demog_disco}
The usual choice of population regulation for population genetics --
a constant size, as in the Wright--Fisher model --
implies biologically unrealistic interactions between geographically distant parts of the species range.
Our choice to regulate population size by including a local density-dependent control on mortality
is only one of many possible ways to do this.
We could have instead regulated fecundity, or recruitment, or both;
this general class of models
is sometimes referred to as the ``Bolker--Pacala model'' \citep{bolker1997using}.
It is not currently clear how much different choices of demographic parameters,
or of functional forms for the regulation,
might quantitatively affect our results,
although the general predictions should be robust to similar forms of regulation.
Since populations are still entirely \emph{intrinsically} regulated,
our model still has a very strong ``population genetics'' flavor.
Alternatively, population size could be regulated by interactions with other species
(e.g., a Lotka-Volterra model),
or extrinsically specified by local resource availability
(e.g., by food or nest site availability).
Indeed, our model could be interpreted as a caricature of such a model:
as local density increases, good habitat is increasingly occupied,
pushing individuals into more marginal habitat and increasing their mortality.
Many such models should behave similarly to ours,
but others (especially those with local population cycling), may differ dramatically.
There is a clear need for greater exploration of 
the consequences for population genetics of ecologically realistic population models.
